### 2023-08-22 resilience.tex
up

In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe,\Bad)$ is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp. a path of length smaller than or equal to $k$) that reaches a state in $\Safe$. We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp. $\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.  $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$). In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.

@@

\problemx{Resilience (RP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
{$\Bad \longrightarrow^{*} \Safe$ ?\newline}

@@

\problemx{$k$-resilience (kRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}

\problemx{Bounded resilience (BRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that {$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}}

@@

\iffalse

\problemx{Resilience problems}
{A transition system $\mathscr{S}=(S,\rightarrow)$, two sets $\Safe, \Bad \subseteq S$, $k \in \mathbb{N}$.}
{ ({\sc resilience problem (RP)}) $\Bad \longrightarrow^{*} \Safe$ ?\newline  
({\sc $k$-resilience problem (kRP)})		$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline
({\sc bounded resilience problem (BRP)})	$\exists k \geq 0$ such that $\mathscr{S}$ is 
 $k$-resilient ?\newline}

\fi

\iffalse
In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
and $\Bad$ is the complement of $\Safe$. 
% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
% wheher
The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.
\fi

\iffalse
\begin{example}
Let us consider a VASS with only one state $q$, one counter, and one transition that substracts $-1$ from the counter, with $\Safe = \downarrow q(0)$, $\Bad = \uparrow q(1)$,
then resilience hold but not bounded resilience: for every bound $k$ there is an element of $\Bad$ e.g. $q(k+1)$ which necessitate at least $k+1$ steps before it can reach $\Safe$. 
\end{example}
\fi

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

It is useful to note that in a sequence of states% of $S$
, every prefix is always 
either in the case that the latest visited $\Bad$ state has already been followed by a 
$\Safe$ state \-- and thus the latest state is contextually ``$\Safe$'' \--
or in the case that it still has to \-- and thus the latest state is contextually ``$\Bad$''. 
Thus, \emph{in context}, there is no true ``neutral'' state.
Hence we will consider in the following not only that $\Bad$ and $\Safe$ are disjoint but
% w.l.o.g.
	also
that $\Bad$ is the complement of $\Safe$,
in which case
if $\Bad$ is downward-closed then $\Safe$ is upward-closed and vice-versa.
Cases where $\Bad$ or $\Safe$ are not complements of each other, and share closure properties, i.e. are both downward-closed or both upward-closed,
have also been studied, 
see 
Appendix~\ref{case down down}
and
Appendix~\ref{case up up} for results.

>>>>

In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe)$ is to decide whether from \emph{any} state in %$\Bad$
$S$, \emph{there exists} a path (resp. a path of length smaller than or equal to $k$) that reaches a state in $\Safe$. We use the notation 
% $\Bad \longrightarrow^{*} \Safe$ (resp. $\Bad \longrightarrow^{\leq k} \Safe$)
$S \longrightarrow^{*} \Safe$ (resp. $S \longrightarrow^{\leq k} \Safe$)
 for 
 %$\forall x \in \Bad, \exists y \in \Safe$ 
$\forall x \in S, \exists y \in \Safe$ 
 such that $x \longrightarrow^{*} y$ 
 %(resp.  $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
  (resp.  $\forall x \in S, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$).
  In our framework, $\Safe \subseteq S$  is possibly infinite but  must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.

@@

\problemx{Resilience (RP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
{$S \longrightarrow^{*} \Safe$ ?\newline}

@@

\problemx{$k$-resilience (kRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and a sets $\Safe \subseteq S$.}
{$S \longrightarrow^{\leq k} \Safe$ ?\newline}

\problemx{Bounded resilience (BRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and a set $\Safe \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that {$S \longrightarrow^{\leq k} \Safe$ ?\newline}}

@@

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$% and $\Bad$
. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@
